---
title: "cleaning_and_exploratory"
title_project: "Cancer epidemiology and "
format: html
editor: visual
---

# References

-   Profiling global cancer incidence and mortality by socioeconomic development (doi:10.1002/ijc.33114) Here, 175 countries were included and 27 cancer types were evaluated

-   Current and future burden of breast cancer: Global statistics for 2020 and 2040 https://doi.org/10.1016/j.breast.2022.08.010 - 185 countries or territories were included - by sex and 18 age groups

-   Global burden of female breast cancer and its association with socioeconomic development status, 1990–2044 - 204 countries - Exposure variables: Metabolic factors and SDI

-   Global and regional trends in incidence and mortality of female breast cancer and associated factors at national level in 2000 to 2019 (10.1097/CM9.0000000000001814)

# Load packages

```{r include = FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(rstatix)
library(ggpubr)
```

# Import data

```{r include=FALSE}
globocan_mortality <- 
  read_csv("~/globocan_and_factors/raw_data/globocan_mortality.csv")

globocan_incidence <- 
  read_csv("~/globocan_and_factors/raw_data/globocan_incidence.csv")

cancer_code =
  read.csv("~/globocan_and_factors/raw_data/cancer_code.csv", 
           stringsAsFactors = T)

hdi_and_comp_2022 = 
  read.csv("~/globocan_and_factors/raw_data/hdi_and_composites_2022.csv")

hdi_and_comp_2017 = 
  read.csv("~/globocan_and_factors/raw_data/hdi_and_composites_2017.csv")

sdi_2018_to_2021 = 
  read.csv("~/globocan_and_factors/raw_data/sdi_2018_to_2021.csv")

regiones_mundo = 
  read.delim("~/globocan_and_factors/raw_data/regiones_mundo.txt", 
             stringsAsFactors = TRUE)
```

# Function to calculate Education and Income (EdI) Index

Estimation of EdI was based on the study of Lortet-Tieulent et al. (2020). For each country, the function below (calculate_EdI) takes as imput: - Expected Years of Scholarity (EYS) - Median Years of Scholarity (MYS) - Gross National Income per capita (GNI)

The function calculate_EdI asummes a max_EYS = 20, max_MYS = 15, min_GNI = 100, and max_GNI = 75000

```{r}
# Define the function to calculate EdI
calculate_EdI = function(EYS, MYS, GNI_per_capita, 
                          max_EYS = 20, max_MYS = 15, 
                          min_GNI = 100, max_GNI = 75000) {
  
  # Normalize education indicators
  norm_EYS <- EYS / max_EYS
  norm_MYS <- MYS / max_MYS
  education_index <- (norm_EYS + norm_MYS) / 2
  
  # Normalize GNI per capita
  norm_GNI <- (log(GNI_per_capita) - log(min_GNI)) / (log(max_GNI) - log(min_GNI))
  
  # Calculate EdI
  EdI <- sqrt(education_index * norm_GNI)
  
  return(EdI)
}
```

# Tidying data

## Epidemiological data

Overall, 185 countries have incidence and mortality data

```{r include=FALSE}
globocan = clean_names(rbind(globocan_incidence, globocan_mortality))
cancer_code = clean_names(cancer_code)


globocan_0 <- merge(globocan, cancer_code, by = "cancer_code", all.x = TRUE)

globocan_0 = 
  globocan_0 |>
  rename(cancer_type = "label.y",
        Country_name = "label.x") |>
  filter(Country_name != "Total") 

countries_with_epid_data = unique(globocan_0$Country_name)
```

## Socioeconomic indicators

Korea (Democratic People's Rep. of) and Monaco were removed because there were no data available regarding HDI and HDI's composites.

The following countries or regions are considered as Developing: Arab States, East Asia and the Pacific, Europe and Central Asia, Latin America and the Caribbean, South Asia, Sub-Saharan Africa.

Cabo verde, France, and Netherlands (as appeared on hdi_and_com_2022) have been changed to Cape verde, France (metropolitans), and The Netherlands (Metropolitan)

Total == World Global estimation

## Calculation and categorization of Education and Income (EdI) Index

```{r}
# Apply the calculate_EdI function to the dataframe
hdi_and_comp_2022_0 = 
  hdi_and_comp_2022 |>
  mutate(Education_and_Income = 
           calculate_EdI(MYS = Mean_years_of_schooling_y,
                         EYS = Expected_years_of_schooling_y,
                         GNI_per_capita = GNI_per_capita_2017PPP_dolar),
         edi_categories = case_when(Education_and_Income < .53 ~ "Low",
                                    Education_and_Income >= .53 &
                                      Education_and_Income < .70 ~ "Medium", 
                                    Education_and_Income > .70 &
                                      Education_and_Income < .80 ~ "High",
                                    Education_and_Income >= .80 ~ "Very high"))
```

## Exploring the number of countries with socioeconomic indicators

Socio Demographic index (SDI), n = 170

Human Development Index (HDI), n = 175

Education and Income (EDI), Index =

```{r}
dim(hdi_and_comp_2022) # [1] 200  10
dim(hdi_and_comp_2017) # [1] 196  10
dim(sdi_2018_to_2021) # [1] 229   5

sdi_2018_to_2021_0 =
  sdi_2018_to_2021 |> 
  filter(sdi_2018_to_2021$Country %in% countries_with_epid_data) ##

hdi_and_comp_2017_0 = 
  hdi_and_comp_2017 |>
  filter(hdi_and_comp_2017$Country %in% countries_with_epid_data)

hdi_and_comp_2022_0 = 
  hdi_and_comp_2022_0 |>
  filter(hdi_and_comp_2022_0$Country %in% countries_with_epid_data) |> 
  rename(Country_name = "Country")

dim(hdi_and_comp_2017_0) # [1] 171  10
dim(hdi_and_comp_2022_0) # [1] 176  10
dim(sdi_2018_to_2021_0) # [1] 170   5

# Countries without indicator data
setdiff(countries_with_epid_data, sdi_2018_to_2021_0$Country)
```

## Matching epidemiological data and indicators

```{r}
sdi_2018_to_2021_0 = 
  sdi_2018_to_2021_0 |> 
  rename(Country_name = "Country")

globocan_1 = globocan_0 %>%
  left_join(hdi_and_comp_2022_0, by = "Country_name")

globocan_2 = 
  globocan_1 |> 
  left_join(sdi_2018_to_2021_0, by = "Country_name")

colnames(globocan_2) = gsub("\\_y$", "", colnames(globocan_2))
globocan_2 = clean_names(globocan_2)
```

## Classifying cancers

```{r}
globocan_3 <- 
  globocan_2 |>
  mutate(
    neoplasia_sex = case_when(
      cancer_type == "Lip, oral cavity" ~ "Both",
      cancer_type == "Salivary glands" ~ "Both",
      cancer_type == "Oropharynx" ~ "Both",
      cancer_type == "Nasopharynx" ~ "Both",
      cancer_type == "Hypopharynx" ~ "Both",
      cancer_type == "Oesophagus" ~ "Both",
      cancer_type == "Stomach" ~ "Both",
      cancer_type == "Colon" ~ "Both",
      cancer_type == "Rectum" ~ "Both",
      cancer_type == "Anus" ~ "Both",
      cancer_type == "Liver and intrahepatic bile ducts" ~ "Both",
      cancer_type == "Gallbladder" ~ "Both",
      cancer_type == "Pancreas" ~ "Both",
      cancer_type == "Larynx" ~ "Both",
      cancer_type == "Trachea, bronchus and lung" ~ "Both",
      cancer_type == "Melanoma of skin" ~ "Both",
      cancer_type == "Non-melanoma skin cancer" ~ "Both",
      cancer_type == "Mesothelioma" ~ "Both",
      cancer_type == "Kaposi sarcoma" ~ "Both",
      cancer_type == "Breast" ~ "Both",
      cancer_type == "Vulva" ~ "Female",
      cancer_type == "Vagina" ~ "Female",
      cancer_type == "Cervix uteri" ~ "Female",
      cancer_type == "Ovary" ~ "Female",
      cancer_type == "Corpus uteri" ~ "Female",
      cancer_type == "Penis" ~ "Male",
      cancer_type == "Prostate" ~ "Male",
      cancer_type == "Testis" ~ "Male",
      cancer_type == "Kidney" ~ "Both",
      cancer_type == "Bladder" ~ "Both",
      cancer_type == "Thyroid" ~ "Both",
      cancer_type == "Hodgkin lymphoma" ~ "Both",
      cancer_type == "Non-Hodgkin lymphoma" ~ "Both",
      cancer_type == "Multiple myeloma" ~ "Both",
      cancer_type == "Leukaemia" ~ "Both",
      cancer_type == "All cancers excl. non-melanoma skin cancer" ~ "Both",
      cancer_type == "Colorectum" ~ "Both",
      cancer_type == " Brain, central nervous system" ~ "Both"),
    tumor_types = case_when(
      cancer_type == "Lip, oral cavity" ~ "Squamous cell carcinoma",
      cancer_type == "Salivary glands" ~ "Adenocarcinoma",
      cancer_type == "Oropharynx" ~ "Squamous cell carcinoma",
      cancer_type == "Nasopharynx" ~ "Squamous cell carcinoma",
      cancer_type == "Hypopharynx" ~ "Squamous cell carcinoma",
      cancer_type == "Oesophagus" ~ "Squamous cell/Adenocarcinoma",
      cancer_type == "Stomach" ~ "Adenocarcinoma",
      cancer_type == "Colon" ~ "Adenocarcinoma",
      cancer_type == "Colorectum" ~ "Adenocarcinoma",
      cancer_type == "Rectum" ~ "Adenocarcinoma",
      cancer_type == "Anus" ~ "Squamous cell/Adenocarcinoma",
      cancer_type == "Liver and intrahepatic bile ducts" ~ 
        "Hepatocellular/Adenocarcinoma",
      cancer_type == "Gallbladder" ~ "Squamous cell/Adenocarcinoma",
      cancer_type == "Pancreas" ~ "Adenocarcinoma",
      cancer_type == "Larynx" ~ "Squamous cell carcinoma",
      cancer_type == "Trachea, bronchus and lung" ~ "Squamous cell/Adenocarcinoma",
      cancer_type == "Melanoma of skin" ~ "Melanoma",
      cancer_type == "Non-melanoma skin cancer" ~ "Squamous/basal cell carcinoma",
      cancer_type == "Mesothelioma" ~ "Mesothelioma",
      cancer_type == "Kaposi sarcoma" ~ "Sarcoma",
      cancer_type == "Breast" ~ "Adenocarcinoma",
      cancer_type == "Vulva" ~ "Squamous cell carcinoma",
      cancer_type == "Vagina" ~ "Squamous cell carcinoma",
      cancer_type == "Cervix uteri" ~ "Squamous cell/Adenocarcinoma",
      cancer_type == "Ovary" ~ "Adenocarcinoma",
      cancer_type == "Corpus uteri" ~ "Adenocarcinoma",
      cancer_type == "Penis" ~ "Squamous cell carcinoma",
      cancer_type == "Prostate" ~ "Adenocarcinoma",
      cancer_type == "Testis" ~ "Adenocarcinoma",
      cancer_type == "Kidney" ~ "Adenocarcinoma",
      cancer_type == "Bladder" ~ "Urothelial carcinoma",
      cancer_type == "Thyroid" ~ "Adenocarcinoma",
      cancer_type == "Hodgkin lymphoma" ~ "Lymphoma",
      cancer_type == "Non-Hodgkin lymphoma" ~ "Lymphoma",
      cancer_type == "Multiple myeloma" ~ "Myeloma",
      cancer_type == "Leukaemia" ~ "Leukaemia",
      cancer_type == "All cancers excl. non-melanoma skin cancer" ~ "Others",
      cancer_type == " Brain, central nervous system" ~ "Brain and spinal cord cancers"),
    malignant_neoplasms_types = case_when(
      cancer_type == "Lip, oral cavity" ~ "Lip, oral cavity and pharynx",
      cancer_type == "Salivary glands" ~ "Lip, oral cavity and pharynx",
      cancer_type == "Oropharynx" ~ "Lip, oral cavity and pharynx",
      cancer_type == "Nasopharynx" ~ "Lip, oral cavity and pharynx",
      cancer_type == "Hypopharynx" ~ "Lip, oral cavity and pharynx",
      cancer_type == "Oesophagus" ~ "Digestive organs",
      cancer_type == "Stomach" ~ "Digestive organs",
      cancer_type == "Colon" ~ "Digestive organs",
      cancer_type == "Colorectum" ~ "Digestive organs",
      cancer_type == "Rectum" ~ "Digestive organs",
      cancer_type == "Anus" ~ "Digestive organs",
      cancer_type == "Liver and intrahepatic bile ducts" ~ "Digestive organs",
      cancer_type == "Gallbladder" ~ "Digestive organs",
      cancer_type == "Pancreas" ~ "Digestive organs",
      cancer_type == "Larynx" ~ "Respiratory and intrathoracic organs",
      cancer_type == "Trachea, bronchus and lung" ~ "Respiratory and intrathoracic organs",
      cancer_type == "Melanoma of skin" ~ "Melanoma and other malignant neoplasms of skin",
      cancer_type == "Non-melanoma skin cancer" ~ "Melanoma and other malignant neoplasms of skin",
      cancer_type == "Mesothelioma" ~ "Mesothelial and soft tissue",
      cancer_type == "Kaposi sarcoma" ~ "Mesothelial and soft tissue",
      cancer_type == "Breast" ~ "Breast",
      cancer_type == "Vulva" ~ "Female genital organs",
      cancer_type == "Vagina" ~ "Female genital organs",
      cancer_type == "Cervix uteri" ~ "Female genital organs",
      cancer_type == "Ovary" ~ "Female genital organs",
      cancer_type == "Corpus uteri" ~ "Female genital organs",
      cancer_type == "Penis" ~ "Male genital organs",
      cancer_type == "Prostate" ~ "Male genital organs",
      cancer_type == "Testis" ~ "Male genital organs",
      cancer_type == "Kidney" ~ "Urinary tract",
      cancer_type == "Bladder" ~ "Urinary tract",
      cancer_type == "Thyroid" ~ "Thyroid and other endocrine glands",
      cancer_type == "Hodgkin lymphoma" ~ "Lymphoid, haematopoietic and related tissue",
      cancer_type == "Non-Hodgkin lymphoma" ~ "Lymphoid, haematopoietic and related tissue",
      cancer_type == "Multiple myeloma" ~ "Lymphoid, haematopoietic and related tissue",
      cancer_type == "Leukaemia" ~ "Lymphoid, haematopoietic and related tissue",
      cancer_type == "All cancers excl. non-melanoma skin cancer" ~ "Other of skin and independent multiple sites",
      cancer_type == " Brain, central nervous system" ~ "Eye, brain and other parts of central nervous system"),
    number_quartiles = case_when(
      indicator == "incidence" & number <= 31.00 ~ "incidence_Q1",
      indicator == "incidence" & number > 31.00 & number <= 195.00 ~ "incidence_Q2",
      indicator == "incidence" & number > 195.00 & number <= 1088.25 ~ "incidence_Q3",
      indicator == "incidence" & number > 1088.25 & number <= 4775419.00 ~ "incidence_Q4",
      indicator == "mortality" & number <= 19.00 ~ "mortality_Q1",
      indicator == "mortality" & number > 19.00 & number <= 112.00 ~ "mortality_Q2",
      indicator == "mortality" & number > 112.00 & number <= 635.75 ~ "mortality_Q3",
      indicator == "mortality" & number > 635.75 & number <= 2560612.00 ~ "mortality_Q4"),
    asr_world_quartiles = case_when(
      indicator == "incidence" & asr_world <= 0.65 ~ "incidence_Q1",
      indicator == "incidence" & asr_world > 0.65 & asr_world <= 2.50 ~ "incidence_Q2",
      indicator == "incidence" & asr_world > 2.50 & asr_world <= 7.20 ~ "incidence_Q3",
      indicator == "incidence" & asr_world > 7.20 & asr_world <= 349.80 ~ "incidence_Q4",
      indicator == "mortality" & asr_world <= 0.31 ~ "mortality_Q1",
      indicator == "mortality" & asr_world > 0.31 & asr_world <= 1.50 ~ "mortality_Q2",
      indicator == "mortality" & asr_world > 1.50 & asr_world <= 4.10 ~ "mortality_Q3",
      indicator == "mortality" & asr_world > 4.10 & asr_world <= 181.30 ~ "mortality_Q4"),
    crude_rate_quartiles = case_when(
      indicator == "incidence" & crude_rate <= 0.64 ~ "incidence_Q1",
      indicator == "incidence" & crude_rate > 0.64 & crude_rate <= 2.40 ~ "incidence_Q2",
      indicator == "incidence" & crude_rate > 2.40 & crude_rate <= 9.60 ~ "incidence_Q3",
      indicator == "incidence" & crude_rate > 9.60 & crude_rate <= 790.60 ~ "incidence_Q4",
      indicator == "mortality" & crude_rate <= 0.33 ~ "mortality_Q1",
      indicator == "mortality" & crude_rate > 0.33 & crude_rate <= 1.40 ~ "mortality_Q2",
      indicator == "mortality" & crude_rate > 1.40 & crude_rate <= 5.90 ~ "mortality_Q3",
      indicator == "mortality" & crude_rate > 5.90 & crude_rate <= 338.40 ~ "mortality_Q4")
    )
```

```{r}
globocan_3 = globocan_3 |> 
  left_join(regiones_mundo, by = "country_name")

```

# Exploratory data analysis

## Globocan

## Descriptive statistics by indicator

```{r}
globocan_2 |>
  distinct(country_name, .keep_all = TRUE) |> 
  filter(indicator == "incidence") |> # Change as appropiate
  select(asr_world, hdi_category,
         education_and_income, human_development_index,
         sdi_2021) |> 
  skim()
```

## Visual inspection

### HDI and composites

```{r}
hdi_counts <- 
  globocan_2 |> 
  filter(hdi_category != "") |> 
  distinct(country_name, .keep_all = TRUE) |> 
  count(hdi_category)

edi_counts = 
  globocan_2 |> 
  filter(edi_categories != "") |> 
  distinct(country_name, .keep_all = TRUE) |> 
  count(edi_categories)

hdi_labels <- data.frame(hdi_category = c("High", "Low", "Medium", "Very high"), 
                       label = c("n = 42", "n = 33", "n = 38", "n = 62"))

edi_labels <- data.frame(edi_categories = c("High", "Low", "Medium", "Very high"), 
                       label = c("n = 34", "n = 43", "n = 49", "n = 49"))
```

```{r}
globocan_2 |> 
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = human_development_index,
             fill = hdi_category)) +
  geom_histogram(alpha = 0.5) +
  facet_wrap(~hdi_category, scales = "free") +
  theme_minimal() +
  geom_text(aes(label = label), data = hdi_labels,
              x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
            inherit.aes = FALSE)
```

```{r}
globocan_2 |>
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = education_and_income,
             fill = edi_categories)) +
  geom_histogram(alpha = 0.5) +
  facet_wrap(~edi_categories, scales = "free") +
  theme_minimal() +
  geom_text(aes(label = label), data = edi_labels,
              x = Inf, y = Inf, hjust=1.1, vjust=1.1,
            inherit.aes = FALSE)
```

```{r}
hdi_p =
  globocan_2 |> 
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = human_development_index,
             fill = hdi_category)) +
  geom_histogram(position = "identity", 
                 alpha = .5) +
  labs(x = "Human Development Index") +
  theme_minimal()

edi_p = 
  globocan_2 |> 
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = education_and_income,
             fill = edi_categories)) +
  geom_histogram(position = "identity", 
                 alpha = .5) +
  labs(x = "Education and Income Index") +
  theme_minimal()

ggarrange(hdi_p, edi_p, nrow = 2, ncol = 1)
```

```{r}
hdi_cat = 
  globocan_2 |>
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = hdi_category, fill = hdi_category)) + 
  geom_bar(alpha = 0.4) +
  theme_minimal()

hdi_prop = 
  globocan_2 |>
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  count(hdi_category) |> 
  mutate(proportion = n / sum(n))  |> 
  ggplot(aes(x = hdi_category, y = proportion,
             fill = hdi_category)) +
  geom_bar(stat = "identity", alpha = 0.4) +
  theme_minimal()

edi_cat = 
  globocan_2 |>
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = edi_categories, fill = edi_categories)) + 
  geom_bar(alpha = .4)

edi_prop = 
  globocan_2 |> 
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  count(edi_categories) |> 
  mutate(proportion = n / sum(n))  |> 
  ggplot(aes(x = edi_categories, y = proportion,
             fill = edi_categories)) +
  geom_bar(stat = "identity", alpha = 0.4) 

ggarrange(hdi_cat, hdi_prop, 
          edi_cat, edi_prop,
          ncol = 2,
          nrow = 2)
```

### Globocan

```{r}
# Histogram
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |> 
  filter(indicator == "incidence") |> 
  ggplot(aes(asr_world)) + 
  geom_histogram() +
  facet_wrap(~cancer_type,
             scales = "free")

globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |> 
  filter(indicator == "mortality") |> 
  ggplot(aes(asr_world)) + 
  geom_histogram() +
  facet_wrap(~cancer_type,
             scales = "free")
```

### Scatterplots

```{r}
# Incidence + mortality vs edi
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |> 
  ggplot(aes(education_and_income, asr_world, color = indicator)) +
  geom_point() +
  facet_wrap(~cancer_type, 
             scales = "free")

valid = c("Breast", "Colon", "Prostate")

# Incidence vs edi
globocan_2 |>
  filter(indicator == "incidence" & cancer_type %in% valid) |> 
  ggplot(aes(education_and_income, asr_world, color = cancer_type)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")

# Incidence + mortality vs hdi
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |> 
  ggplot(aes(human_development_index, asr_world, color = indicator)) +
  geom_point() +
  facet_wrap(~indicator, 
             scales = "free")

globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |> 
  ggplot(aes(human_development_index, asr_world)) +
  geom_point() +
  facet_wrap(~indicator, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")

# Incidence vs EdI by tumor type
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |>
  filter(indicator == "incidence") |> 
  ggplot(aes(education_and_income, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized incidence rate") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           #color = "cornflowerblue", 
           geom = "label")

# Mortality vs EdI by tumor type
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |>
  filter(indicator == "mortality") |> 
  ggplot(aes(education_and_income, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized mortality rate") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           #color = "cornflowerblue", 
           geom = "label")


# Incidence vs human_development_index by tumor type
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |>
  filter(indicator == "incidence") |> 
  ggplot(aes(human_development_index, asr_world)) +
  geom_point() +
  labs(y = "Age-standardized incidence rate",
       x = "Human development index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")

# Mortality vs human_development_index by tumor type
globocan_2 |>
  distinct(cancer_type, country_name, .keep_all = TRUE) |>
  filter(indicator == "mortality") |> 
  ggplot(aes(human_development_index, asr_world)) +
  geom_point() +
  labs(y = "Age-standardized mortality rate",
       x = "Human development index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")

```

# Model 3. Generalised additive models

- Main references: 10.1097/CM9.0000000000001814 (Global and regional trends in incidence and mortality of female breast cancer and associated factors at national level in 2000 to 2019). Chinese journal

- https://bookdown.org/ssjackson300/Machine-Learning-Lecture-Notes/generalised-additive-models.html. Background

# Model 4. anova on selected cancer: those with normal distribution
Step 1: select cancer types
Step 2: categorize HDI on 5, for comparison (https://bmjopen.bmj.com/content/bmjopen/9/10/e028461.full.pdf)


