---
title: "cleaning_and_exploratory"
title_project: "Cancer epidemiology and "
format: html
editor: visual
---
## Literature

Profiling global cancer incidence and mortality by socioeconomic 
development (doi:10.1002/ijc.33114)
  - 175 countries were included
  - 27 cancer types were evaluated

Current and future burden of breast cancer: Global statistics for 2020 and 2040
https://doi.org/10.1016/j.breast.2022.08.010
  - 185 countries or territories were included
  - by sex and 18 age groups

Global burden of female breast cancer and its association with socioeconomic development status, 1990–2044
  - 204 countries
  - Exposure variables: Metabolic factors and SDI

## Load packages

```{r include = FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(rstatix)
```

## Import data

```{r include = FALSE}
## Incidence and cancer mortality
globocan_mortality <- read_csv("globocan_mortality.csv")
globocan_incidence <- read_csv("globocan_incidence.csv")

## Socioeconomic indicators
hdi_and_comp_2022 = read.csv("~/globocan_and_factors/raw_data/hdi_and_composites_2022.csv")
hdi_and_comp_2017 = read.csv("~/globocan_and_factors/raw_data/hdi_and_composites_2017.csv")
sdi_2018_to_2021 = read.csv("~/globocan_and_factors/raw_data/SDI_2018to2021.csv")

```

## Tidying data

### Cancer epidemiology
```{r}
globocan = rbind(globocan_incidence, globocan_mortality)
#globocan$Country <- NULL
globocan = 
  globocan |> rename(Country = Label)
globocan$`Cancer code` = as.character(globocan$`Cancer code`)
```

### Socioeconomic indicators
Korea (Democratic People's Rep. of) and Monaco were removed because there was no
data available regarding HDI and HDI's composites.

The following countries or regions are considered as Developing: Arab States, 
East Asia and the Pacific, Europe and Central Asia, Latin America and the Caribbean, 
South Asia, Sub-Saharan Africa.

Cabo verde, France, and Netherlands (as appeared on hdi_andcom_2022) have been changed to 
Cape verde, France, and The Netherlands (Metropolitan)

Total = World
Global estimation

#### Filtering countries with both indicator and epidemiological data
```{r}
dim(hdi_and_comp_2022) # [1] 194  10
dim(hdi_and_comp_2017) # [1] 196  10
dim(sdi_2018_to_2021) # [1] 229   5

sdi_2018_to_2021 =
  sdi_2018_to_2021 |> rename(Country = Location)  # [1] 229 5

sdi_218_to_2021_0 =
  sdi_2018_to_2021 |> 
  filter(sdi_2018_to_2021$Country %in% countries_with_epid_data) ## [1] 159   5

hdi_and_comp_2017_0 = 
  hdi_and_comp_2017 |>
  filter(hdi_and_comp_2017$Country %in% countries_with_epid_data)

hdi_and_comp_2022_0 = 
  hdi_and_comp_2022 |>
  filter(hdi_and_comp_2022$Country %in% countries_with_epid_data)

dim(hdi_and_comp_2017_0) # [1] 171  10
dim(hdi_and_comp_2022_0) # [1] 176  10
dim(sdi_218_to_2021_0) # [1] 158   5

# Countries without indicator data
setdiff(countries_with_epid_data, sdi_218_to_2021_0$Country) ## correct the names !
```
## Exploration of the data

### Globocan
```{r}
dim(globocan) ## [1] 14136 observations
length(unique(globocan$Label)) ## [1] 186 countries
length(unique(globocan$`Cancer code`)) ## [1] 39 types of cancer

dim(globocan |> filter(indicator == "incidence")) ## [1] 7068   11
dim(globocan |> filter(indicator == "mortality")) ## [1] 7068   11
skim(globocan)
```

### HDI and composites

```{r}
skim(hdi_edi_and_comp_2022)

hdi_edi_and_comp_2022  |>  
  count(hdi_category) |> 
  adorn_totals()  
```
## Visual inspection

### HDI and composites
```{r}
globocan |>
  ggplot(aes(x=`ASR (World)`)) +
  geom_histogram() +
  xlim(0, 100) +
  facet_wrap(~indicator, scales = "free")

hdi_edi_and_comp_2022 |> ## We need another categorical variable
  count(hdi_category, Human_Development_Index) |> 
  ggplot() +
    geom_col(
      mapping = aes(   
        x = hdi_category,
        fill = hdi_category,         
        y = n))
```
### Globocan

```{r}
globocan |> ## We need another categorical variable
  count(hdi_category, Human_Development_Index) |> 
  ggplot() +
    geom_col(
      mapping = aes(   
        x = hdi_category,
        fill = hdi_category,         
        y = n))

globocan |>
  filter(indicator == "incidence") |> 
  group_by(`Cancer code`) |>
  mutate( 
    quatile = rstatix::get_summary_stats(Number, type = "quantile"))
  
```


## Function to calculate Education and Income (EdI) Index

It is based on the study of Lortet-Tieulent et al. (2020). 
Set values
test 2. 
Step 1.

test 1. max_EYS = 20, max_MYS = 15, min_GNI = 100, max_GNI = 75000
EdI Panama = 0.7225345, 
EdI Panama = 0.742 (reference)

test 2. max_EYS = 20, max_MYS = 15, min_GNI = 100, max_GNI = 75000

```{r}
# Define the function to calculate EdI
calculate_EdI <- function(EYS, MYS, GNI_per_capita, 
                          max_EYS = 20, max_MYS = 15, 
                          min_GNI = 100, max_GNI = 75000) {
  
  # Normalize education indicators
  norm_EYS <- EYS / max_EYS
  norm_MYS <- MYS / max_MYS
  education_index <- (norm_EYS + norm_MYS) / 2
  
  # Normalize GNI per capita
  norm_GNI <- (log(GNI_per_capita) - log(min_GNI)) / (log(max_GNI) - log(min_GNI))
  
  # Calculate EdI
  EdI <- sqrt(education_index * norm_GNI)
  
  return(EdI)
}
```

## Calculate EdI
```{r}
# Apply the calculate_EdI function to the dataframe
hdi_edi_and_comp_2022 <- hdi_and_comp_2022 |> 
  mutate(EdI = calculate_EdI(MYS = Mean_years_of_schooling_y,
                             EYS = Expected_years_of_schooling_y,
                             GNI_per_capita = GNI_per_capita_2017PPP_dolar))

# Print the dataframe with the new EdI column
print(data)
```

