---
title: "cleaning_and_exploratory"
title_project: "Cancer epidemiology and "
format: html
editor: source
---

# Main references

-   Profiling global cancer incidence and mortality by socioeconomic development (doi:10.1002/ijc.33114). Here 175 countries were included and 27 cancer types were evaluated

-   Current and future burden of breast cancer: Global statistics for 2020 and 2040 https://doi.org/10.1016/j.breast.2022.08.010. 185 countries or territories were included - by sex and 18 age groups

-   Global burden of female breast cancer and its association with socioeconomic development status, 1990–2044 - 204 countries - Exposure variables: Metabolic factors and SDI

-   Global and regional trends in incidence and mortality of female breast cancer and associated factors at national level in 2000 to 2019 (10.1097/CM9.0000000000001814)

# Load packages \[-\]

```{r include = FALSE}
pacman::p_load(
  tidyverse,
  skimr, 
  janitor,
  rstatix,
  ggpubr,
  broom,
  car,
  tidymodels,
  performance,
  rio, 
  here,
  janitor,
  splines,
  ggformula,
  mgcv)
```

# Import data

Incidence and mortality were retrieved from GLOBOCAN (<https://gco.iarc.fr/>)

Human Development Index (hdi), hdi composites, and socio demographic (sdi) index were retrieved from (Junior completalo !!!)

SDI and quintiles were retrieved from IHME-Global Health Exchange Data: https://ghdx.healthdata.org/record/global-burden-disease-study-2021-gbd-2021-socio-demographic-index-sdi-1950%E2%80%932021

```{r include=FALSE}
# Epidemiology
globocan_mortality <-
  import(here("data", "raw", "globocan_mortality.csv"))

globocan_incidence <-
  import(here("data", "raw", "globocan_incidence.csv"))

miratio <- globocan_mortality$`ASR (World)` / globocan_incidence$`ASR (World)`

miratio_df <- globocan_incidence %>%
  mutate(`ASR (World)` = miratio,
         indicator = 'miratio') 

globocan_incidence_and_miratio <- bind_rows(globocan_incidence, miratio_df)

# Sociodemographic indicators
hdi_and_comp_2022 <-
  import(here("data", "raw", "hdi_and_composites_2022.csv"))

hdi_and_comp_2017 <-
  import(here("data", "raw", "hdi_and_composites_2017.csv"))

sdi_2018_to_2021 <-
  import(here("data", "raw", "sdi_2018_to_2021.csv"))

gdb_sdi_2021_and_quintiles <-
  import(here("data", "raw", "ihme_gdb_sdi_2021_and_quintiles.csv"))

# Complementary data
regiones_mundo <- import(here("data", "raw", "regiones_mundo.txt"))

cancer_code <- import(here("data", "raw", "cancer_code.csv"))
```

# Function to calculate Education and Income (EdI) Index

Estimation of EdI was based on the study of Lortet-Tieulent et al. (2020) (https://doi.org/10.1002/ijc.33114). For each country, the function below (calculate_EdI) takes as imput: - Expected Years of Scholarity (EYS) - Median Years of Scholarity (MYS) - Gross National Income per capita (GNI)

The function calculate_EdI asummes a max_EYS = 20, max_MYS = 15, min_GNI = 100, and max_GNI = 75000

Imput data were retrieved from (Junior completalo)

```{r}
# Define the function to calculate EdI
calculate_EdI = function(EYS, MYS, GNI_per_capita, 
                          max_EYS = 20, max_MYS = 15, 
                          min_GNI = 100, max_GNI = 75000) {
  
  # Normalize education indicators
  norm_EYS <- EYS / max_EYS
  norm_MYS <- MYS / max_MYS
  education_index <- (norm_EYS + norm_MYS) / 2
  
  # Normalize GNI per capita
  norm_GNI <- (log(GNI_per_capita) - log(min_GNI)) / (log(max_GNI) - log(min_GNI))
  
  # Calculate EdI
  EdI <- sqrt(education_index * norm_GNI)
  
  return(EdI)
}
```

## Calculation and categorization of Education and Income (EdI) Index

```{r}
# Apply the calculate_EdI function to the dataframe
hdi_and_comp_2022_0 = 
  hdi_and_comp_2022 |>
  mutate(Education_and_Income = 
           calculate_EdI(MYS = Mean_years_of_schooling_y,
                         EYS = Expected_years_of_schooling_y,
                         GNI_per_capita = GNI_per_capita_2017PPP_dolar),
         edi_categories = case_when(Education_and_Income < .53 ~ "Low",
                                    Education_and_Income >= .53 &
                                      Education_and_Income < .70 ~ "Medium", 
                                    Education_and_Income >= .70 &
                                      Education_and_Income < .80 ~ "High",
                                    Education_and_Income >= .80 ~ "Very high"))
```

# Tidying data

## Epidemiological data

Overall, 185 countries have incidence and mortality data

```{r include=FALSE}
globocan <- janitor::clean_names(rbind(globocan_incidence_and_miratio, 
                                       globocan_mortality))
cancer_code <- janitor::clean_names(cancer_code)

globocan_0 <- merge(globocan, cancer_code, by = "cancer_code", all.x = TRUE)

globocan_0 <- globocan_0 |>
  dplyr::rename(cancer_type = "label.y", Country_name = "label.x") |>
  dplyr::filter(Country_name != "Total")

countries_with_epid_data <- unique(globocan_0$Country_name)
countries_with_epid_data = unique(globocan_0$Country_name)
```

## Socioeconomic indicators

Korea (Democratic People's Rep. of) and Monaco were removed because there were no data available regarding **hdi** and **hdi**'s composites.

The following countries or regions are considered as Developing countries: Arab States, East Asia and the Pacific, Europe and Central Asia, Latin America and the Caribbean, South Asia, Sub-Saharan Africa.

Cabo verde, France, and Netherlands (as appeared on hdi_and_com_2022) have been changed to Cape verde, France (metropolitans), and The Netherlands (Metropolitan) Venezuela (Boliarian Republic of) has been changed to Venezuela France to France (metropolitan) Democratic People's Republic of Korea has been changed to Korea, Democratic People Republic of Iran (Islamic Republic of) to Iran, Islamic Republic of Democratic Republic of the Congo to Congo, Democratic People Republic of Gambia to The Republic of the Gambia Netherlands to The Netherlands Republic of Korea to Korea, Republic of Cabo Verde to Cape Verde Republic of Moldova to Moldova Bosnia and Herzegovina to Bosnia Herzegovina Bolivia (Plurinational State of) to Bolivia United Republic of Tanzania to Tanzania, United Republic of China (without Hong Kong and Macao) to China

Total == World Global estimation

## Exploring the number of countries with socioeconomic indicators

Socio Demographic index (SDI), n = 170 Human Development Index (HDI), n = 175 Education and Income (EDI), Index =

```{r}
#SDI
sdi_2018_to_2021_0 =
  sdi_2018_to_2021 |> 
  filter(sdi_2018_to_2021$Country %in% countries_with_epid_data) ##

gdb_sdi_2021_and_quintiles = 
  gdb_sdi_2021_and_quintiles |> 
  janitor::clean_names()

gdb_sdi_2021_and_quintiles_0 =
  gdb_sdi_2021_and_quintiles |> 
  select(-location_id) |>
  rename(Country_name = "location_name",
         sdi_gdb_2021 = "x2021_sdi_index_value",
         sdi_categories = "sdi_quintile") |> 
  dplyr::filter(gdb_sdi_2021_and_quintiles$location_name %in% countries_with_epid_data)

#HDI
hdi_and_comp_2017_0 = 
  hdi_and_comp_2017 |>
  filter(hdi_and_comp_2017$Country %in% countries_with_epid_data)

hdi_and_comp_2022_0 = 
  hdi_and_comp_2022_0 |>
  filter(hdi_and_comp_2022_0$Country %in% countries_with_epid_data) |> 
  rename(Country_name = "Country")

# Countries without indicator data. Change as appropriate
setdiff(countries_with_epid_data, gdb_sdi_2021_and_quintiles$Location.Name)
```

## Matching epidemiological data and indicators

```{r}
sdi_2018_to_2021_0 = 
  sdi_2018_to_2021_0 |> 
  rename(Country_name = "Country")

globocan_1 = globocan_0 %>%
  left_join(hdi_and_comp_2022_0, by = "Country_name")

globocan_2 = 
  globocan_1 |> 
  left_join(sdi_2018_to_2021_0, by = "Country_name")

globocan_3 =
  globocan_2 |> 
  left_join(gdb_sdi_2021_and_quintiles_0, by = "Country_name")

colnames(globocan_3) = gsub("\\_y$", "", colnames(globocan_3))
globocan_3 = clean_names(globocan_3)
```

## Classifying cancers

```{r}
globocan_4 <- globocan_3 |>
  dplyr::mutate(
    neoplasia_sex = case_when(
      cancer_type %in% c(
        "Lip, oral cavity", "Salivary glands", "Oropharynx", "Nasopharynx",
        "Hypopharynx", "Oesophagus", "Stomach", "Colon", "Rectum", "Anus",
        "Liver and intrahepatic bile ducts", "Gallbladder", "Pancreas",
        "Larynx", "Trachea, bronchus and lung", "Melanoma of skin",
        "Non-melanoma skin cancer", "Mesothelioma", "Kaposi sarcoma",
        "Breast", "Kidney", "Bladder", "Thyroid", "Hodgkin lymphoma",
        "Non-Hodgkin lymphoma", "Multiple myeloma", "Leukaemia",
        "All cancers excl. non-melanoma skin cancer", "Colorectum",
        " Brain, central nervous system") ~ "Both",
      cancer_type %in% c(
        "Vulva", "Vagina", "Cervix uteri", "Ovary", "Corpus uteri") ~ "Female",
      cancer_type %in% c(
        "Penis", "Prostate", "Testis") ~ "Male"
    ),
    tumor_types = case_when(
      cancer_type %in% c(
        "Lip, oral cavity", "Oropharynx", "Nasopharynx", "Hypopharynx", "Larynx",
        "Vulva", "Vagina", "Penis") ~ "Squamous cell carcinoma",
      cancer_type %in% c(
        "Salivary glands", "Stomach", "Colon", "Colorectum", "Rectum", "Pancreas",
        "Breast", "Ovary", "Corpus uteri", "Prostate", "Testis", "Kidney",
        "Thyroid") ~ "Adenocarcinoma",
      cancer_type %in% c(
        "Gallbladder", "Oesophagus", "Anus", "Trachea, bronchus and lung",
        "Cervix uteri") ~ "Squamous cell/Adenocarcinoma",
      cancer_type %in% c("Hodgkin lymphoma", "Non-Hodgkin lymphoma") ~ "Lymphoma",
      cancer_type %in% c("Melanoma of skin", "Non-melanoma skin cancer") ~ "Skin",
      TRUE ~ "Others"
    ), 
    malignant_neoplasms_types = case_when(
      cancer_type %in% c(
        "Lip, oral cavity", "Salivary glands", "Oropharynx", "Nasopharynx", 
        "Hypopharynx") ~ "Lip, oral cavity and pharynx",
      cancer_type %in% c(
        "Oesophagus", "Stomach", "Colon", "Colorectum", "Rectum", "Anus",
        "Liver and intrahepatic bile ducts", "Gallbladder", 
        "Pancreas") ~ "Digestive organs",
      cancer_type %in% c(
        "Larynx", "Trachea, bronchus and lung") ~ "Respiratory and intrathoracic organs",
      cancer_type %in% c(
        "Melanoma of skin",
        "Non-melanoma skin cancer") ~ "Melanoma and other malignant neoplasms of skin",
      cancer_type %in% c(
        "Mesothelioma", "Kaposi sarcoma") ~ "Mesothelial and soft tissue",
      cancer_type == "Breast" ~ "Breast",
      cancer_type %in% c(
        "Vulva", "Vagina", "Cervix uteri", "Ovary", 
        "Corpus uteri") ~ "Female genital organs",
      cancer_type %in% c("Penis", "Prostate", "Testis") ~ "Male genital organs",
      cancer_type %in% c("Kidney", "Bladder") ~ "Urinary tract",
      cancer_type %in% c(
        "Hodgkin lymphoma", "Non-Hodgkin lymphoma", "Multiple myeloma",
        "Leukaemia") ~ "Lymphoid, haematopoietic and related tissue",
      cancer_type == "Thyroid" ~ "Thyroid and other endocrine glands",
      cancer_type == " Brain, central nervous system" ~ "Eye, brain and other parts of central nervous system",
      TRUE ~ "Others"
    ), 
    number_quartiles = case_when(
      indicator == "incidence" & number <= 31.00 ~ "incidence_Q1",
      indicator == "incidence" &
        number > 31.00 & number <= 195.00 ~ "incidence_Q2",
      indicator == "incidence" &
        number > 195.00 & number <= 1088.25 ~ "incidence_Q3",
      indicator == "incidence" &
        number > 1088.25 & number <= 4775419.00 ~ "incidence_Q4",
      indicator == "mortality" & number <= 19.00 ~ "mortality_Q1",
      indicator == "mortality" &
        number > 19.00 & number <= 112.00 ~ "mortality_Q2",
      indicator == "mortality" &
        number > 112.00 & number <= 635.75 ~ "mortality_Q3",
      indicator == "mortality" &
        number > 635.75 & number <= 2560612.00 ~ "mortality_Q4"
    ),
    asr_world_quartiles = case_when(
      indicator == "incidence" & asr_world <= 0.65 ~ "incidence_Q1",
      indicator == "incidence" &
        asr_world > 0.65 & asr_world <= 2.50 ~ "incidence_Q2",
      indicator == "incidence" &
        asr_world > 2.50 & asr_world <= 7.20 ~ "incidence_Q3",
      indicator == "incidence" &
        asr_world > 7.20 & asr_world <= 349.80 ~ "incidence_Q4",
      indicator == "mortality" & asr_world <= 0.31 ~ "mortality_Q1",
      indicator == "mortality" &
        asr_world > 0.31 & asr_world <= 1.50 ~ "mortality_Q2",
      indicator == "mortality" &
        asr_world > 1.50 & asr_world <= 4.10 ~ "mortality_Q3",
      indicator == "mortality" &
        asr_world > 4.10 & asr_world <= 181.30 ~ "mortality_Q4"
    ),
    crude_rate_quartiles = case_when(
      indicator == "incidence" & crude_rate <= 0.64 ~ "incidence_Q1",
      indicator == "incidence" &
        crude_rate > 0.64 & crude_rate <= 2.40 ~ "incidence_Q2",
      indicator == "incidence" &
        crude_rate > 2.40 & crude_rate <= 9.60 ~ "incidence_Q3",
      indicator == "incidence" &
        crude_rate > 9.60 & crude_rate <= 790.60 ~ "incidence_Q4",
      indicator == "mortality" &
        crude_rate <= 0.33 ~ "mortality_Q1",
      indicator == "mortality" &
        crude_rate > 0.33 & crude_rate <= 1.40 ~ "mortality_Q2",
      indicator == "mortality" &
        crude_rate > 1.40 & crude_rate <= 5.90 ~ "mortality_Q3",
      indicator == "mortality" &
        crude_rate > 5.90 & crude_rate <= 338.40 ~ "mortality_Q4"
    )
  )
```

### Adding world regions

```{r}
globocan_5 = globocan_4 |> 
  dplyr::left_join(regiones_mundo, by = "country_name")
```

```{r}
globocan_6 <- globocan_5 |>
  dplyr::select(
    indicator,
    cancer_type,
    country_name,
    asr_world,
    crude_rate,
    human_development_index,
    education_and_income,
    sdi_gdb_2021,
    hdi_category,
    edi_categories,
    sdi_categories,
    neoplasia_sex
  ) |>
  dplyr::rename(
    `ASR (World)` = asr_world,
    `Crude rate` = crude_rate,
    Indicator = indicator,
    HDI = human_development_index,
    EdI = education_and_income,
    SDI = sdi_gdb_2021
  ) |>
  dplyr::mutate(
    Indicator = factor(Indicator) |>
      forcats::fct_recode(
        "Incidence" = "incidence", 
        "Mortality" = "mortality") |>
      forcats::fct_relevel("Incidence", "Mortality"),
    country_label = case_when(
      country_name %in% c(
        "United States of America", "Bolivia", "Spain", "Ethiopia", "China",
        "Japan", "India", "Peru", "Egypt", "Saudi Arabia", 
        "France (metropolitan)", "Uganda") ~ as.character(country_name),
      TRUE ~ NA_character_
    ) |>
      forcats::fct_recode(
        "USA" = "United States of America", "France" = "France (metropolitan)")
  )

# Complex names data
globocan_7 <- globocan_6 |>
  dplyr::distinct(cancer_type, country_name, Indicator, .keep_all = TRUE)

# Clean names data
globocan_8 <- globocan_7 |>
  janitor::clean_names() |>
  dplyr::mutate(
    cancer_type = case_when(
      cancer_type == " Brain, central nervous system" ~ "Brain, nervous system",
      cancer_type == "Liver and intrahepatic bile ducts" ~ "Liver",
      .default = as.character(cancer_type))
  )
```

# Exploratory data analysis

## Globocan

## Descriptive statistics by indicator

```{r}
globocan_8 |>
  distinct(country_name, 
           cancer_type, 
           indicator, .keep_all = TRUE) |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  filter(indicator == "miratio" & cancer_type == "Breast") |> # Change as appropiate
  select(asr_world) |> 
  skim()
```

## Visual inspection

### HDI and composites

```{r}
hdi_counts <- 
  globocan_8 |> 
  filter(hdi_category != "") |> 
  distinct(country_name, .keep_all = TRUE) |> 
  count(hdi_category)

edi_counts = 
  globocan_8 |> 
  filter(edi_categories != "") |> 
  distinct(country_name, .keep_all = TRUE) |> 
  count(edi_categories)

hdi_labels <- data.frame(hdi_category = c("High", "Low", "Medium", "Very high"), 
                       label = c("n = 42", "n = 33", "n = 38", "n = 62"))

edi_labels <- data.frame(edi_categories = c("High", "Low", "Medium", "Very high"), 
                       label = c("n = 34", "n = 43", "n = 49", "n = 49"))
```

```{r}
globocan_8 |> 
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = human_development_index,
             fill = hdi_category)) +
  geom_histogram(alpha = 0.5) +
  facet_wrap(~hdi_category, scales = "free") +
  theme_minimal() +
  geom_text(aes(label = label), data = hdi_labels,
              x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
            inherit.aes = FALSE, 
            size = 7.5) +
  theme(text = element_text(size = 20))
```

```{r}
globocan_8 |>
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = education_and_income,
             fill = edi_categories)) +
  geom_histogram(alpha = 0.5) +
  facet_wrap(~edi_categories, scales = "free") +
  theme_minimal() +
  geom_text(aes(label = label), data = edi_labels,
              x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
            inherit.aes = FALSE,
            size = 7.5) +
  theme(text = element_text(size = 20))
```

```{r}
hdi_p =
  globocan_5 |> 
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = human_development_index,
             fill = hdi_category)) +
  geom_histogram(position = "identity", 
                 alpha = .5) +
  labs(x = "Human Development Index") +
  theme_minimal() +
  theme(text = element_text(size = 20))

edi_p = 
  globocan_5 |> 
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = education_and_income,
             fill = edi_categories)) +
  geom_histogram(position = "identity", 
                 alpha = .5) +
  labs(x = "Education and Income Index") +
  theme_minimal() +
  theme(text = element_text(size = 20))

ggarrange(hdi_p, edi_p, nrow = 2, ncol = 1)

sdi_p = 
  globocan_5 |> 
  filter(sdi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = sdi_gdb_2021,
             fill = sdi_categories)) +
  geom_histogram(position = "identity", 
                 alpha = .5) +
  labs(x = "Sociodemographic index") +
  theme_minimal() +
  theme(text = element_text(size = 20))

ggarrange(hdi_p, edi_p, sdi_p,
          nrow = 3, ncol = 1)
```

### Proportions and counts by categories

```{r}
hdi_cat = 
  globocan_5 |>
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = hdi_category, fill = hdi_category)) + 
  geom_bar(alpha = 0.4) +
  theme_minimal()

hdi_prop = 
  globocan_5 |>
  filter(hdi_category != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  count(hdi_category) |> 
  mutate(proportion = n / sum(n))  |> 
  ggplot(aes(x = hdi_category, y = proportion,
             fill = hdi_category)) +
  geom_bar(stat = "identity", alpha = 0.4) +
  theme_minimal()

edi_cat = 
  globocan_5 |>
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = edi_categories, fill = edi_categories)) + 
  geom_bar(alpha = .4)

edi_prop = 
  globocan_5 |> 
  filter(edi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  count(edi_categories) |> 
  mutate(proportion = n / sum(n))  |> 
  ggplot(aes(x = edi_categories, y = proportion,
             fill = edi_categories)) +
  geom_bar(stat = "identity", alpha = 0.4) 

sdi_cat = 
  globocan_5 |>
  filter(sdi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  ggplot(aes(x = sdi_categories, fill = sdi_categories)) + 
  geom_bar(alpha = .4)

sdi_prop = 
  globocan_5 |> 
  filter(sdi_categories != "") |>
  distinct(country_name, .keep_all = TRUE) |> 
  count(sdi_categories) |> 
  mutate(proportion = n / sum(n))  |> 
  ggplot(aes(x = sdi_categories, y = proportion,
             fill = sdi_categories)) +
  geom_bar(stat = "identity", alpha = 0.4) 


ggarrange(hdi_cat, hdi_prop, 
          edi_cat, edi_prop,
          sdi_cat, sdi_prop,
          ncol = 2,
          nrow = 3)
```

### Histogram of the distribution of epidemiological indicators

```{r}
mean_incidence <- globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Incidence") |>
  group_by(cancer_type) |>
  summarise(mean_asr_world = mean(asr_world, na.rm = TRUE))

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |> 
  ggplot(aes(asr_world)) + 
  geom_histogram() +
  facet_wrap(~cancer_type,
             scales = "free") +
  geom_vline(data = mean_incidence, 
             aes(xintercept = mean_asr_world), 
             color = "salmon", 
             linetype = "dashed", 
             size = 2) +
  labs(x = "Count", 
       y = "Age-standarized incidence rate")


mean_mortality = 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Mortality") |>
  group_by(cancer_type) |>
  summarise(mean_asr_world = mean(asr_world, na.rm = TRUE))

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Mortality") |> 
  ggplot(aes(asr_world)) + 
  geom_histogram() +
  facet_wrap(~cancer_type,
             scales = "free") +
  geom_vline(data = mean_mortality, 
             aes(xintercept = mean_asr_world), 
             color = "salmon", 
             linetype = "dashed", 
             size = 1) +
  labs(x = "Count", 
       y = "Age-standarized mortality rate")

mean_miratio = 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "miratio") |>
  group_by(cancer_type) |>
  summarise(mean_asr_world = mean(asr_world, na.rm = TRUE))

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "miratio") |> 
  ggplot(aes(asr_world)) + 
  geom_histogram() +
  facet_wrap(~cancer_type,
             scales = "free") +
  geom_vline(data = mean_miratio, 
             aes(xintercept = mean_asr_world), 
             color = "salmon", 
             linetype = "dashed", 
             size = 1) +
  labs(x = "Count", 
       y = "Incidence to mortality ratio")
```

### qqplots by cancer type

```{r}
globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Mortality") |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "miratio") |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()
```

### Scatterplots epidemiological vs socioeconomic indicator

Epidemiological indicator vs EdI by tumor type

```{r}
p_incidence_v_edi =
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Incidence") |> 
  ggplot(aes(ed_i, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized incidence rate") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")
```

```{r}
p_mortality_v_edi = 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Mortality") |> 
  ggplot(aes(ed_i, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized mortality rate") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "salmon", 
           geom = "label")
```

```{r}
p_miratio_v_edi = 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "miratio") |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(ed_i, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized mortality rate") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "black", 
           geom = "label")
```

Epidemiological indicator vs hdi by tumor type

```{r}
p_incidence_v_hdi = 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Incidence") |> 
  ggplot(aes(hdi, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standarized mortality rate") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")
```

```{r}
p_mortality_v_hdi = 
  globocan_8 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Mortality") |> 
  ggplot(aes(hdi, asr_world)) +
  geom_point() +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized mortality rate",
       x = "Human development index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "salmon", 
           geom = "label")
```

```{r}
p_miratio_v_hdi = 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "miratio") |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(hdi, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Mortality to incidence ratio",
       x = "Human development index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "black", 
           geom = "label")
```

Epidemiological indicator vs sdi by tumor type

```{r}
p_incidence_v_sdi = 
  globocan_8 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Incidence") |> 
  ggplot(aes(sdi, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized incidence rate",
       x = "Sociodemographic index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "cornflowerblue", 
           geom = "label")
```

```{r}
p_motality_v_sdi = 
  globocan_5 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "Mortality") |> 
  ggplot(aes(sdi, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Age-standardized mortality rate",
       x = "Sociodemographic index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "salmon", 
           geom = "label")
```

```{r}
p_miratio_v_sdi = 
  globocan_8 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |>
  filter(indicator == "miratio") |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sdi, asr_world)) +
  geom_point(alpha = 0.5) +
  labs(y = "Mortality to incidence ratio",
       x = "Sociodemographic index") +
  facet_wrap(~cancer_type, 
             scales = "free") +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(aes(label = after_stat(rr.label)), 
           color = "black", 
           geom = "label")
```

# Model 1. Simple linear model grouped by cancer_type

https://stats.stackexchange.com/questions/551264/when-to-check-model-assumptions

https://stackoverflow.com/questions/1169539/linear-regression-and-group-by-in-r

### Incidence

```{r}
models_incidence_edi <- globocan_8 |>
  dplyr::filter(indicator == "Incidence") |>
  dplyr::group_by(cancer_type) |>
  tidyr::nest() |>
  dplyr::mutate(model = purrr::map(data, ~ lm(asr_world ~ ed_i, data = .)))

# Summarizes information about the models
m1_incidence_edi <- models_incidence_edi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_incidence_edi <- models_incidence_edi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_EdI_incidence <- dplyr::bind_cols(m1_incidence_edi, m2_incidence_edi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

```{r}
models_incidence_hdi <- globocan_8 |>
  dplyr::filter(indicator == "Incidence") |>
  dplyr::group_by(cancer_type) |>
  tidyr::nest() |>
  dplyr::mutate(model = purrr::map(data, ~ lm(asr_world ~ hdi, data = .)))

# Summarizes information about the models
m1_incidence_hdi <- models_incidence_hdi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_incidence_hdi <- models_incidence_hdi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_HDI_incidence <- dplyr::bind_cols(m1_incidence_hdi, m2_incidence_hdi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

```{r}
models_incidence_sdi <- globocan_8 |>
  dplyr::filter(indicator == "Incidence") |>
  dplyr::group_by(cancer_type) |>
  tidyr::nest() |>
  dplyr::mutate(model = purrr::map(data, ~ lm(asr_world ~ sdi, data = .)))

# Summarizes information about the models
m1_incidence_sdi <- models_incidence_sdi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_incidence_sdi <- models_incidence_sdi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_SDI_incidence <- dplyr::bind_cols(m1_incidence_sdi, m2_incidence_sdi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

### Mortality

```{r}
models_mortality_edi <- globocan_8 |>
  dplyr::filter(indicator == "Mortality") |>
  dplyr::group_by(cancer_type) |>
  tidyr::nest() |>
  dplyr::mutate(model = purrr::map(data, ~ lm(asr_world ~ ed_i, data = .)))

# Summarizes information about the models
m1_mortality_edi <- models_mortality_edi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_mortality_edi <- models_mortality_edi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_EdI_mortality <- dplyr::bind_cols(m1_mortality_edi, m2_mortality_edi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

```{r}
models_mortality_hdi <- globocan_8 |>
  dplyr::filter(indicator == "Mortality") |>
  dplyr::group_by(cancer_type) |>
  tidyr::nest() |>
  dplyr::mutate(model = purrr::map(data, ~ lm(asr_world ~ hdi, data = .)))

# Summarizes information about the models
m1_mortality_hdi <- models_mortality_hdi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_mortality_hdi <- models_mortality_hdi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_HDI_mortality <- dplyr::bind_cols(m1_mortality_hdi, m2_mortality_hdi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

```{r}
models_mortality_sdi <- globocan_8 |>
  dplyr::filter(indicator == "Mortality") |>
  dplyr::group_by(cancer_type) |>
  tidyr::nest() |>
  dplyr::mutate(model = purrr::map(data, ~ lm(asr_world ~ sdi, data = .)))

# Summarizes information about the models
m1_mortality_sdi <- models_mortality_sdi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_mortality_sdi <- models_mortality_sdi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_SDI_mortality <- dplyr::bind_cols(m1_mortality_sdi, m2_mortality_sdi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

### Mortality to incidence ratio

```{r}
models_miratio_edi <- globocan_8 |>
  dplyr::filter(indicator == "miratio") |>
  group_by(cancer_type) |> 
  tidyr::nest() |>
  dplyr::mutate(
    data = purrr::map(data, ~ dplyr::filter(.x, !is.na(asr_world) & !is.na(ed_i) & is.finite(asr_world) & is.finite(ed_i))),
    model = purrr::map(data, ~ lm(asr_world ~ ed_i, data = .))
  )

# Summarizes information about the models
m1_miratio_edi <- models_miratio_edi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 4))

m2_miratio_edi <- models_miratio_edi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 4))

# Renamed
ASR_EdI_miratio <- dplyr::bind_cols(m1_miratio_edi, m2_miratio_edi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

```{r}
models_miratio_hdi <- globocan_8 |>
  dplyr::filter(indicator == "miratio") |>
  group_by(cancer_type) |> 
  tidyr::nest() |>
  dplyr::mutate(
    data = purrr::map(data, ~ dplyr::filter(.x, !is.na(asr_world) & !is.na(hdi) & is.finite(asr_world) & is.finite(hdi))),
    model = purrr::map(data, ~ lm(asr_world ~ hdi, data = .))
  )

# Summarizes information about the models
m1_miratio_hdi <- models_miratio_hdi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 4))

m2_miratio_hdi <- models_miratio_hdi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 4))

# Renamed
ASR_HDI_miratio <- dplyr::bind_cols(m1_miratio_hdi, m2_miratio_hdi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

```{r}
models_miratio_sdi <- globocan_8 |>
  dplyr::filter(indicator == "miratio") |>
  group_by(cancer_type) |> 
  tidyr::nest() |>
  dplyr::mutate(
    data = purrr::map(data, ~ dplyr::filter(.x, !is.na(asr_world) & !is.na(sdi) & is.finite(asr_world) & is.finite(sdi))),
    model = purrr::map(data, ~ lm(asr_world ~ sdi, data = .))
  )

# Summarizes information about the models
m1_miratio_sdi <- models_miratio_sdi |>
  dplyr::mutate(summary = map(model, broom::tidy)) |>
  tidyr::unnest(summary) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::select(-data, -model, -term) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

m2_miratio_sdi <- models_miratio_sdi |>
  dplyr::mutate(summary = map(model, broom::glance)) |>
  tidyr::unnest(summary) |>
  dplyr::select(-data, -model) |>
  dplyr::mutate(across(where(is.numeric), round, 2))

# Renamed
ASR_SDI_miratio <- dplyr::bind_cols(m1_miratio_sdi, m2_miratio_sdi) |>
  dplyr::select(
    cancer_type...1,
    estimate,
    r.squared,
    adj.r.squared,
    statistic...10,
    p.value...11,
    AIC,
    BIC,
    deviance
  ) |>
  dplyr::rename(
    `Cancer type` = cancer_type...1,
    Estimate = estimate,
    `R-squared` = r.squared,
    `Adjusted R-squared` = adj.r.squared,
    `F-statistic` = statistic...10,
    `p-value` = p.value...11,
    Deviance = deviance
  )
```

# Model 2. One-factor ANOVA on selected cancers

Main Reference: https://bmjopen.bmj.com/content/bmjopen/9/10/e028461.full.pdf

Categorical variables: hdi, sdi, edi Continuous variables: Mortality, Incidence

Pairwise comparisons Tamhane T2 test for variables with normal distribution but heterogeneous variances ------- for

### For Education and Income index (EDI) vs Incidence

Visual inspection

```{r}
p_incidence_v_edi

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |>
  #filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()
```

Perform ANOVA, extract residuals, test normality and homogeneity of variances by cancer type

```{r include=FALSE}
stat_incid_v_edi <-
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |>
  group_by(cancer_type) |>
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ edi_categories, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ edi_categories, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_incid_v_edi)


```

**Table 1.** Selection of 5 cancer types based on normality test, qqplot

| N°  | Cancer type                   | R\^2 | Shapiro p value | Levene p value | anova p values |
|-----|-------------------------------|------|-----------------|----------------|----------------|
| 1   | Breast                        |      | 0.358           | \<0.001        | \<0.001        |
| 2   | Hodgkin lymphoma              |      | 0.558           | \<0.001        | \<0.001        |
| 3   | Leukaemia                     |      | 0.251           | 0.002          | \<0.001        |
| 4   | Non-hodgkin lymphoma          |      | 0.819           | \<0.001        | \<0.001        |
| 5   | Brain, central nervous system |      | 0.635           | 0.001          | \<0.001        |

Pairwise comparison for normally distributed data with unequal variances

```{r}
pw_gh_test_inc_edi = 
  globocan_8 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence" & 
           cancer_type %in% c("Breast", 
                              "Hodgkin lymphoma",
                              "Non-Hodgkin lymphoma",
                              "Leukaemia",
                              "Brain, nervous system")) |> 
  group_by(cancer_type) |> 
  rstatix::games_howell_test(asr_world ~ edi_categories)

View(pw_gh_test_inc_edi)
```

### Education and Income index (EDI) vs mortality

Visual inspection

```{r}
p_mortality_v_edi

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Mortality") |>
  #filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()
```

Perform ANOVA, extract residuals, test normality and homogeneity of variances by cancer type

```{r include=FALSE}
stat_mort_v_edi <- 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Mortality") |>
  group_by(cancer_type) |>
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ edi_categories, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ edi_categories, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_mort_v_edi)
```

**Table 2.** Selection of cancer types and statistics for mortality vs edi

| N°  | Cancer type                   | R\^2 | Shapiro P value | Levene P value | ANOVA P values |     |
|-----|-------------------------------|------|-----------------|----------------|----------------|-----|
| 1   | Brain, central nervous system |      | 0.158           | 0.004          | \<0.001        |     |

Pairwise comparison for normally distributed data with unequal variances

```{r}
pw_gh_test_mort_hdi = 
  globocan_8 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Mortality" & 
           cancer_type %in% c("Brain, nervous system")) |> 
  group_by(cancer_type) |> 
  rstatix::games_howell_test(asr_world ~ edi_categories)
View(pw_gh_test_mort_hdi)
```

### Mortality to incidence ratio vs Education and Income index (EDI)

Visual inspection

```{r}
p_miratio_v_edi

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "miratio") |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()
```

Perform ANOVA, extract residuals, test normality and homogeneity of variances by cancer type

```{r include=FALSE}
stat_miratio_v_edi <- 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "miratio") |>
  filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  group_by(cancer_type) |> 
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ edi_categories, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ edi_categories, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_miratio_v_edi)
```

**Note:** No cancer type showed a normal distribution

| N°  | Cancer type | R\^2 | Shapiro P value | Levene P value | ANOVA P values |
|-----|-------------|------|-----------------|----------------|----------------|
|     |             |      |                 |                |                |

Pairwise comparison for normally distributed data with unequal variances

```{r}
# pw_gh_test_mort_hdi = 
#   globocan_5 |> 
#   distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
#   filter(indicator == "mortality" & 
#            cancer_type %in% c(" Brain, central nervous system")) |> 
#   group_by(cancer_type) |> 
#   rstatix::games_howell_test(asr_world ~ edi_categories)
# View(pw_gh_test_mort_hdi)
```

### Human Development Index (HDI) vs Incidence

Visual inspection

```{r}
p_incidence_v_hdi

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |>
  #filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()
```

Perform ANOVA, extract residuals, test normality and homogeneity of variances by cancer type

```{r include=FALSE}
stat_incid_v_hdi <-
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |>
  group_by(cancer_type) |>
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ hdi_category, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ hdi_category, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_incid_v_hdi)
```

**Table 3.** Selection of cancer types and statistics for incidence vs hdi

| N°  | Cancer type                   | R\^2 | Shapiro p value | Levene p value | anova p values |     |
|-----|-------------------------------|------|-----------------|----------------|----------------|-----|
| 1   | Breast                        |      | 0.920           | \<0.001        | \<0.001        |     |
| 2   | Hodking Lymphoma              |      | 0.838           | \<0.001        | \<0.001        |     |
| 3   | Brain, central nervous system |      | 0.562           | 0.001          |                |     |
| 4   | Non-Hodgkin lymphoma          |      | 0.180           | \<0.001        |                |     |

Pairwise comparison for normally distributed data with unequal variances

```{r}
pw_gh_test_inc_hdi = 
  globocan_8 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence" & 
           cancer_type %in% c("Breast", 
                              "Hodgkin lymphoma",
                              "Brain, nervous system",
                              "Non-Hodgkin lymphoma")) |> 
  group_by(cancer_type) |> 
  rstatix::games_howell_test(asr_world ~ hdi_category)
```

### Mortality vs Human Development Index (HDI)

Visual inspection

```{r}
p_incidence_v_hdi

globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Incidence") |>
  #filter(!is.na(asr_world) & !is.nan(asr_world) & !is.infinite(asr_world)) |>
  ggplot(aes(sample = asr_world)) + 
  stat_qq() + 
  stat_qq_line() + 
  facet_wrap(~ cancer_type,
             scales = "free") + 
  labs(x = "Theoretical Quantiles", 
       y = "Sample Quantiles") + 
  theme_minimal()
```

Perform ANOVA, extract residuals, test normality and homogeneity of variances by cancer type

```{r}
stat_mort_v_hdi <- 
  globocan_8 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "Mortality") |>
  group_by(cancer_type) |>
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ hdi_category, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ hdi_category, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_mort_v_hdi)
```

**Table 4.** Selection of cancer types and statistics for mortality vs hdi

| N°  | Cancer type                   | R\^2 | Shapiro p value | Levene p value | anova p values |     |
|-----|-------------------------------|------|-----------------|----------------|----------------|-----|
| 1   | Brain, central nervous system |      | 0.279           | 0.005          | \<0.001        |     |
| 2   | Kidney                        |      | 0.263           | \<0.001        | \<0.001        |     |
| 3   | Colon                         |      | 0.058           | \<0.001        | \<0.001        |     |

Pairwise comparison for normally distributed data with equal variances

```{r}
pw_gh_test_mort_hdi = 
  globocan_5 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "mortality" & 
           cancer_type %in% c("Kidney", 
                              " Brain, central nervous system",
                              "Colon")) |> 
  group_by(cancer_type) |> 
  rstatix::games_howell_test(asr_world ~ hdi_category)
```

### Sociodemographic index (SDI) vs Incidence

Visual inspection

```{r}

```

Perform ANOVA, extract residuals, test normality and homogeneity of variances by cancer type

```{r}
stat_incid_v_sdi <-
  globocan_5 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "incidence") |>
  group_by(cancer_type) |>
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ sdi_categories, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ sdi_categories, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_incid_v_sdi)
```

**Table 5.** Selection of cancer types and statistics for incidence vs sdi

| N°  | Cancer type                   | R\^2 | Shapiro p value | Levene p value | anova p values |     |
|-----|-------------------------------|------|-----------------|----------------|----------------|-----|
| 1   | Hodgkin lymphoma              |      | 0.896           | \<0.001        | \<0.001        |     |
| 2   | Leukaemia                     |      | 0.495           | 0.023          | \<0.001        |     |
| 3   | Brain, central nervous system |      | 0.486           | \<0.001        | \<0.001        |     |
| 4   | Breast                        |      | 0.185           | \<0.001        | \<0.001        |     |
| 5   | Prostate                      |      | 0.119           | 0.034          | \<0.001        |     |

Pairwise comparison for normally distributed data with equal variances

```{r}
pw_gh_test_inc_sdi = 
  globocan_5 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "incidence" & 
           cancer_type %in% c("Hodgkin lymphoma", 
                              "Leukaemia",
                              " Brain, central nervous system",
                              "Breast",
                              "Prostate")) |> 
  group_by(cancer_type) |> 
  mutate(sdi_categories = na_if(sdi_categories, "")) |> 
  rstatix::games_howell_test(asr_world ~ sdi_categories)
```

### Sociodemographic index (SDI) vs Mortality

```{r}
stat_mort_v_sdi <- 
  globocan_5 |>
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "mortality") |>
  group_by(cancer_type) |>
  do({
    # Perform one-factor ANOVA
    anova_model <- aov(asr_world ~ sdi_categories, data = .)
    
    # Extract residuals
    residuals_data <- residuals(anova_model)
    
    # Perform Levene's Test for homogeneity of variances
    levene_test <- leveneTest(asr_world ~ sdi_categories, data = .)
    
    # Extract ANOVA p-value
    anova_summary <- summary(anova_model)
    anova_p_value <- anova_summary[[1]]$`Pr(>F)`[1]
    
    # Check if residuals are identical
    if (length(unique(residuals_data)) == 1) {
      # Return NA for groups with identical residuals
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = NA,
        shapiro_statistic = NA,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    } else {
      # Test normality of residuals
      shapiro_test <- shapiro.test(residuals_data)
      
      # Return results
      data.frame(
        cancer_type = unique(.$cancer_type),
        shapiro_p_value = shapiro_test$p.value,
        shapiro_statistic = shapiro_test$statistic,
        levene_p_value = levene_test$`Pr(>F)`[1],
        anova_p_value = anova_p_value
      )
    }
  }) |>
  ungroup()

View(stat_mort_v_sdi)
```

**Table 6.** Selection of cancer types and statistics for mortality vs sdi

| N°  | Cancer type                   | R\^2 | Shapiro p value | Levene p value | anova p values |     |
|-----|-------------------------------|------|-----------------|----------------|----------------|-----|
| 1   | Brain, central nervous system |      | 0.699           | 0.013          | \<0.001        |     |
| 2   | Colorectum                    |      | 0.311           | \<0.001        | \<0.001        |     |
| 3   | Kidney                        |      | 0.214           | 0.002          | \<0.001        |     |
| 4   | Pancreas                      |      | 0.106           | \<0.001        | \<0.001        |     |

Pairwise comparison for normally distributed data with equal variances

```{r}
pw_gh_test_mort_sdi = 
  globocan_5 |> 
  distinct(cancer_type, country_name, indicator, .keep_all = TRUE) |> 
  filter(indicator == "mortality" & 
           cancer_type %in% c("Kidney", 
                              "Colorectum",
                              " Brain, central nervous system",
                              "Pancreas")) |> 
  group_by(cancer_type) |> 
  rstatix::games_howell_test(asr_world ~ sdi_categories)
```

# Model 2. Logarithmic models

```{r}
model_log <- lm(education_and_income ~ log(asr_world), data = data_test)

b <- ggplot(data_test, aes(y = asr_world, x = education_and_income)) +
  geom_point() +
  geom_smooth(
    method = "lm",
    formula = y ~ log(x),
    se = FALSE,
    col = "dodgerblue1") +
  theme_light() +
  facet_wrap(~cancer_type, 
             scales = "free")
b
```

# Model 3. Polynomial models

```{r fig.width=10}
model_poly_quadratic <- lm(education_and_income ~ asr_world + I(asr_world^2), data = data_test)

c <- ggplot(data_test, aes(y = asr_world, x = education_and_income)) +
  geom_point() +
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x ^ 2), 
    se = FALSE,
    col = "dodgerblue1") +
  theme_light() +
  facet_wrap(~cancer_type, 
             scales = "free")
c
```

```{r}
model_poly_cubic <- lm(education_and_income ~ asr_world + I(asr_world^2) + I(asr_world^3),
                       data = data_test)

d <- ggplot(data_test, aes(y = asr_world, x = education_and_income)) +
  geom_point() +
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x ^ 2) + I(x ^ 3), 
    se = FALSE,
    col = "dodgerblue1") +
  theme_light() +
  facet_wrap(~cancer_type, 
             scales = "free")
d
```

# Model 4. Polynomial models with splines

```{r}
test <- data_test |>
  dplyr::select(education_and_income, asr_world)
plot(test, pch=19, ylim=c(0, 15), las=1)
abline(v=0.7, lty="dotted", col='tomato')
abline(v=0.85, lty="dotted", col='tomato')
text(x=0.7, y=0.3, 't=0.55', col='tomato')
text(x=0.85, y=0.3, 't=0.55', col='tomato')
```

```{r fig.width=10}
model_spline_1_knot <- lm(education_and_income ~ ns(asr_world, df = 3), data_test)

x <- data_test$asr_world
y <- data_test$education_and_income

plot(x, y, main = "Spline Regression", xlab = "X", ylab = "Y")
lines(x, predict(model_spline_1_knot), col = "red", lwd = 2)
legend("topright", legend = "Fitted Spline", col = "red", lwd = 2)

d <- ggplot(data_test, aes(y = asr_world, x = education_and_income)) +
  geom_point()+
  geom_smooth(
    method = "lm",
    formula = y ~ x + I(x ^ 2) + I(x ^ 3), 
    se = FALSE,
    col = "dodgerblue1") + 
  stat_spline() +
  theme_light() +
  facet_wrap(~cancer_type, 
             scales = "free")
d

ggplot(d,aes(x,y)) + geom_point() + 
  geom_line(data=data.frame(spline(d, n=n*10)))

performance::compare_performance(model_poly_cubic, model_spline_1_knot)
broom::glance(model_poly_cubic)
broom::glance(model_spline_1_knot)
```

```{r fig.width=12}
plot(compare_performance(model_poly_cubic, model_spline_1_knot))
```

```{r}
fit.sp = smooth.spline(education_and_income ~ asr_world, nknots=15,  data_test)
```

```{r}
gam(wage ~ s(year, 4) + s(age, 5) + education, data = Wage)
```
